---
title: "Session 4: Homework 2"
author: "Your name goes here"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---


```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, include=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(here)
library(skimr)
library(janitor)
library(httr)
library(readxl)
library(vroom)
```



# Climate change and temperature anomalies 


If we wanted to study climate change, we can find data on the *Combined Land-Surface Air and Sea-Surface Water Temperature Anomalies* in the Northern Hemisphere at [NASA's Goddard Institute for Space Studies](https://data.giss.nasa.gov/gistemp). The [tabular data of temperature anomalies can be found here](https://data.giss.nasa.gov/gistemp/tabledata_v3/NH.Ts+dSST.txt)

To define temperature anomalies you need to have a reference, or base, period which NASA clearly states that it is the period between 1951-1980.

Run the code below to load the file:

```{r weather_data, cache=TRUE}

weather <- 
  read_csv("https://data.giss.nasa.gov/gistemp/tabledata_v3/NH.Ts+dSST.csv", 
           skip = 1, 
           na = "***")

```

Notice that, when using this function, we added two options: `skip` and `na`.

1. The `skip=1` option is there as the real data table only starts in Row 2, so we need to skip one row. 
1. `na = "***"` option informs R how missing observations in the spreadsheet are coded. When looking at the spreadsheet, you can see that missing data is coded as "***". It is best to specify this here, as otherwise some of the data is not recognized as numeric data.

Once the data is loaded, notice that there is a object titled `weather` in the `Environment` panel. If you cannot see the panel (usually on the top-right), go to `Tools` > `Global Options` > `Pane Layout` and tick the checkbox next to `Environment`. Click on the `weather` object, and the dataframe will pop up on a seperate tab. Inspect the dataframe.

For each month and year, the dataframe shows the deviation of temperature from the normal (expected). Further the dataframe is in wide format. 

You have two objectives in this section:

1. Select the year and the twelve month variables from the `weather` dataset. We do not need the others (J-D, D-N, DJF, etc.) for this assignment. Hint: use `select()` function.

1. Convert the dataframe from wide to 'long' format. Hint: use `gather()` or `pivot_longer()` function. Name the new dataframe as `tidyweather`, name the variable containing the name of the month as `month`, and the temperature deviation values as `delta`.


```{r tidyweather}

```

Inspect your dataframe. It should have three variables now, one each for 

1. year, 
1. month, and 
1. delta, or temperature deviation.

## Plotting Information

Let us plot the data using a time-series scatter plot, and add a trendline. To do that, we first need to create a new variable called `date` in order to ensure that the `delta` values are plot chronologically. 


> In the following chunk of code, I used the `eval=FALSE` argument, which does not run a chunk of code; I did so that you can knit the document before tidying the data and creating a new dataframe `tidyweather`. When you actually want to run this code and knit your document, you must delete `eval=FALSE`, **not just here but in all chunks were `eval=FALSE` appears.**


```{r scatter_plot, eval=FALSE}

tidyweather <- tidyweather %>%
  mutate(date = ymd(paste(as.character(Year), Month, "1")),
         month = month(date, label=TRUE),
         year = year(date))

ggplot(tidyweather, aes(x=date, y = delta))+
  geom_point()+
  geom_smooth(color="red") +
  theme_bw() +
  labs (
    title = "Weather Anomalies"
  )

```

Is the effect of increasing temperature more pronounced in some months? Use `facet_wrap()` to produce a seperate scatter plot for each month, again with a smoothing line. Your chart should human-readable labels; that is, each month should be labeled "Jan", "Feb", "Mar" (full or abbreviated month names are fine), not `1`, `2`, `3`. 

```{r facet_wrap, echo=FALSE}

#Your code goes here...

```

